---
title: "Leaflet"
author: "The Johns Hopkins Data Science Lab"
always_allow_html: yes
---

## Introduction

Leaflet is one of the most popular Javascript libraries for
creating interactive maps. The leaflet R package allows you
to create your own leaflet maps without needing to know any
Javascript!

### Installation

```r
install.packages("leaflet")
```

## Your First Map

Getting started with leaflet is easy. The `leaflet()`
function creates a map widget that you can store in a
variable so that you can modify the map later on. You can
add features to the map using the pipe operator (`%>%`) just
like in dplyr. The `addTiles()` function adds mapping data
from [Open Street Map](http://www.openstreetmap.org/).

```{r, eval=FALSE}
library(leaflet)
my_map <- leaflet() %>% 
  addTiles()
my_map
```

## Your First Map

```{r, echo=FALSE}
library(leaflet)
my_map <- leaflet() %>% 
  addTiles()
my_map
```

## Adding Markers

You can add markers to your map one at a time using the
`addMarkers()` function by specifying the longitude and
latitude. ([Here's](https://twitter.com/drob/status/719927537330040832)
a tip if you tend to mix them up.) You can specify popup
text for when you click on the marker with the `popup` 
argument.

```{r, eval=FALSE}
library(leaflet)
my_map <- my_map %>%
  addMarkers(lat=39.2980803, lng=-76.5898801, 
             popup="Jeff Leek's Office")
my_map
```

## Adding Markers

```{r, echo=FALSE}
library(leaflet)
my_map <- my_map %>%
  addMarkers(lat=39.2980803, lng=-76.5898801, 
             popup="Jeff Leek's Office")
my_map
```

## Adding Many Markers

Adding one marker at a time is often not practical if you
want to display many markers. If you have a data frame with
columns `lat` and `lng` you can pipe that data frame into
`leaflet()` to add all the points at once.

```{r, eval=FALSE}
set.seed(2016-04-25)
df <- data.frame(lat = runif(20, min = 39.2, max = 39.3),
                 lng = runif(20, min = -76.6, max = -76.5))
df %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers()
```

## Adding Many Markers

```{r, echo=FALSE}
set.seed(2016-04-25)
df <- data.frame(lat = runif(20, min = 39.2, max = 39.3),
                 lng = runif(20, min = -76.6, max = -76.5))
df %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers()
```

## Making Custom Markers

The blue markers that leaflet comes packaged with may not be
enough depending on what you're mapping. Thankfully you can
make your own markers from `.png` files.

```{r, eval=FALSE}
hopkinsIcon <- makeIcon(
  iconUrl = "http://brand.jhu.edu/content/uploads/2014/06/university.shield.small_.blue_.png",
  iconWidth = 31*215/230, iconHeight = 31,
  iconAnchorX = 31*215/230/2, iconAnchorY = 16
)

hopkinsLatLong <- data.frame(
  lat = c(39.2973166, 39.3288851, 39.2906617),
  lng = c(-76.5929798, -76.6206598, -76.5469683))

hopkinsLatLong %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers(icon = hopkinsIcon)
```

## Making Custom Markers

```{r, echo=FALSE}
hopkinsIcon <- makeIcon(
  iconUrl = "http://brand.jhu.edu/content/uploads/2014/06/university.shield.small_.blue_.png",
  iconWidth = 31*215/230, iconHeight = 31,
  iconAnchorX = 31*215/230/2, iconAnchorY = 16
)

hopkinsLatLong <- data.frame(
  lat = c(39.2973166, 39.3288851, 39.2906617, 39.2970681, 39.2824806),
  lng = c(-76.5929798, -76.6206598, -76.5469683, -76.6150537, -76.6016766))

hopkinsLatLong %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers(icon = hopkinsIcon)
```

## Adding Multiple Popups

When adding multiple markers to a map, you may want to add
popups for each marker. You can specify a string of plain
text for each popup, or you can provide HTML which will be
rendered inside of each popup.

```{r, eval=FALSE}
hopkinsSites <- c(
  "<a href='http://www.jhsph.edu/'>East Baltimore Campus</a>",
  "<a href='https://apply.jhu.edu/visit/homewood/'>Homewood Campus</a>",
  "<a href='http://www.hopkinsmedicine.org/johns_hopkins_bayview/'>Bayview Medical Center</a>",
  "<a href='http://www.peabody.jhu.edu/'>Peabody Institute</a>",
  "<a href='http://carey.jhu.edu/'>Carey Business School</a>"
)

hopkinsLatLong %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers(icon = hopkinsIcon, popup = hopkinsSites)
```

## Adding Multiple Popups

```{r, echo=FALSE}
hopkinsSites <- c(
  "<a href='http://www.jhsph.edu/'>East Baltimore Campus</a>",
  "<a href='https://apply.jhu.edu/visit/homewood/'>Homewood Campus</a>",
  "<a href='http://www.hopkinsmedicine.org/johns_hopkins_bayview/'>Bayview Medical Center</a>",
  "<a href='http://www.peabody.jhu.edu/'>Peabody Institute</a>",
  "<a href='http://carey.jhu.edu/'>Carey Business School</a>"
)

hopkinsLatLong %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers(icon = hopkinsIcon, popup = hopkinsSites)
```

## Mapping Clusters

Sometimes you might have so many points on a map that it
doesn't make sense to plot every marker. In these situations
leaflet allows you to plot clusters of markers using
`addMarkers(clusterOptions = markerClusterOptions())`. When
you zoom in to each cluster, the clusters will separate until
you can see the individual markers.

```{r, eval=FALSE}
df <- data.frame(lat = runif(500, min = 39.25, max = 39.35),
                 lng = runif(500, min = -76.65, max = -76.55))
df %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers(clusterOptions = markerClusterOptions())
```

## Mapping Clusters

```{r, echo=FALSE}
df <- data.frame(lat = runif(500, min = 39.25, max = 39.35),
                 lng = runif(500, min = -76.65, max = -76.55))
df %>% 
  leaflet() %>%
  addTiles() %>%
  addMarkers(clusterOptions = markerClusterOptions())
```

## Mapping Circle Markers

Instead of adding markers or clusters you can easily add
circle markers using `addCircleMarkers()`.

```{r, eval=FALSE}
df <- data.frame(lat = runif(20, min = 39.25, max = 39.35),
                 lng = runif(20, min = -76.65, max = -76.55))
df %>% 
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers()
```

## Mapping Circle Markers

```{r, echo=FALSE}
df <- data.frame(lat = runif(20, min = 39.25, max = 39.35),
                 lng = runif(20, min = -76.65, max = -76.55))
df %>% 
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers()
```

## Drawing Circles

You can draw arbitrary shapes on the maps you create,
including circles and squares. The code below draws a map
where the circle on each city is proportional to the
population of that city.

```{r, eval=FALSE}
md_cities <- data.frame(name = c("Baltimore", "Frederick", "Rockville", "Gaithersburg", 
                                 "Bowie", "Hagerstown", "Annapolis", "College Park", "Salisbury", "Laurel"),
                        pop = c(619493, 66169, 62334, 61045, 55232,
                                39890, 38880, 30587, 30484, 25346),
                        lat = c(39.2920592, 39.4143921, 39.0840, 39.1434, 39.0068, 39.6418, 38.9784, 38.9897, 38.3607, 39.0993),
                        lng = c(-76.6077852, -77.4204875, -77.1528, -77.2014, -76.7791, -77.7200, -76.4922, -76.9378, -75.5994, -76.8483))
md_cities %>%
  leaflet() %>%
  addTiles() %>%
  addCircles(weight = 1, radius = sqrt(md_cities$pop) * 30)
```

## Drawing Circles

```{r, echo=FALSE}
md_cities <- data.frame(name = c("Baltimore", "Frederick", "Rockville", "Gaithersburg", 
                                 "Bowie", "Hagerstown", "Annapolis", "College Park", "Salisbury", "Laurel"),
                        pop = c(619493, 66169, 62334, 61045, 55232,
                                39890, 38880, 30587, 30484, 25346),
                        lat = c(39.2920592, 39.4143921, 39.0840, 39.1434, 39.0068, 39.6418, 38.9784, 38.9897, 38.3607, 39.0993),
                        lng = c(-76.6077852, -77.4204875, -77.1528, -77.2014, -76.7791, -77.7200, -76.4922, -76.9378, -75.5994, -76.8483))
md_cities %>%
  leaflet() %>%
  addTiles() %>%
  addCircles(weight = 1, radius = sqrt(md_cities$pop) * 30)
```

## Drawing Rectangles

You can add rectangles on leaflet maps as well:

```{r, eval=FALSE}
leaflet() %>%
  addTiles() %>%
  addRectangles(lat1 = 37.3858, lng1 = -122.0595, 
                lat2 = 37.3890, lng2 = -122.0625)
```

## Drawing Rectangles

```{r, echo=FALSE}
leaflet() %>%
  addTiles() %>%
  addRectangles(lat1 = 37.3858, lng1 = -122.0595, 
                lat2 = 37.3890, lng2 = -122.0625)
```

## Adding Legends

Adding a legend can be useful if you have markers on your
map with different colors:

```{r, eval=FALSE}
df <- data.frame(lat = runif(20, min = 39.25, max = 39.35),
                 lng = runif(20, min = -76.65, max = -76.55),
                 col = sample(c("red", "blue", "green"), 20, replace = TRUE),
                 stringsAsFactors = FALSE)

df %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(color = df$col) %>%
  addLegend(labels = LETTERS[1:3], colors = c("blue", "red", "green"))
```

## Adding Legends

```{r, echo=FALSE, message=FALSE}
df <- data.frame(lat = runif(20, min = 39.25, max = 39.35),
                 lng = runif(20, min = -76.65, max = -76.55),
                 col = sample(c("red", "blue", "green"), 20, replace = TRUE),
                 stringsAsFactors = FALSE)

df %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(color = df$col) %>%
  addLegend(labels = LETTERS[1:3], colors = c("blue", "red", "green"))
```

## Conclusion

For more details about the leaflet package for R 
visit http://rstudio.github.io/leaflet/.
